This thesis would not exist without the support I have received from people in the Princeton community and beyond. I am indebted to my advisor, Professor Michael E. Mueller, for his generosity with guidance and encouragement. I thank him for the myriad of ways he has contributed to my intellectual and professional growth.

I am beholden to my graduate student peers and Professors Elie Bou-Zeid, Luigi Martinelli, and Howard A. Stone for assisting me with general exam preparations. I thank Professors Marcus Hultmark, Yiguang Ju, Chung K. Law, and Alexander Smits for being my examiners. I would also like to express my gratitude to my advisor and Professor Yiguang Ju for serving as my reading committee.

Many aspects of this thesis would not have come to fruition without the cooperation of various outstanding researchers outside Princeton University. I am fortunate to have collaborated with Professor Antonio Attili and Mr. Lukas K. W. Berger at RWTH Aachen University and the following researchers at the University of Adelaide: Mr. Saleh Mahmoud, Professor Zeyad T. Alwahabi, Professor Bassam B. Dally, and Professor Graham J. Nathan.

I am grateful for the companionship of my colleagues in CTRFL: Dr. Temistocle Grenga, Dr. Suo Yang, Dr. Carla Bahri, Dr. Sili Deng, Jonathan F. MacArt, A. Cody Nunno, Sandra Sowah, Bruce A. Perry, and Alex G. Novoselov. The discussions that I had with them helped me overcome obstacles in my research, and their presence in lab reminded me that it is always sunny in D101.

I extend my appreciation to Seongwoo Oh, Justin L. Ripley, Jonathan Shi, and James E. Park for being compassionate cohabitants, and I am thankful for the Big Red comradeship of Seongwoo Oh, Zhaoqi Leng, Anthony Savas, and Daniel Floryan. Finally, I deeply thank my parents for their steadfast encouragement and support.

This work would not be possible without the valuable support in the form of computational time on the TIGRESS high performance computer center at Princeton University, which is jointly supported by the Princeton Institute for Computational Science and Engineering (PICSciE) and the Princeton University Office of Information Technology's Research Computing Department. This thesis carries T-3342 in the records of the Department of Mechanical and Aerospace Engineering.

%% Advisor
%% Thesis committee/generals committee
%% Professors who helped me during the general exam process
%% labmates + other labs + classmates + Daniel floryan
%% Jill Ray and Theresa
%% Susanne Killian
%% Roommates Seong, Justin
%% Ian, Bryan, going to the gym
%% Collaborators such as Bassam Dally
%% Susanne Killian and Shalin Shah
%% Advisors from Cornell (Olivier Desjardins, Elizabeth Fisher, Dr. Brandon Hencey)
%% Parents

%% I would like to thank the Math department for providing the original documentclass file that this class is based upon. I would like to thank my parents, without whom my life would not be possible. I would also like to thank my advisor, my dissertation committee, and my research collaborators because every graduate student needs to do so. And finally, I thank the members of my research group, to whom I leave this template to save you some of the trouble I had to go through getting my dissertation to compile in \LaTeX{}.  

%% Don't forget to ask your advisor if your work was sponsored by a grant that needs to be acknowledged in this section.

%% Lastly, I want to thank my parents for their 
