\chapter{Modeling Framework for Large Eddy Simulation\label{ch:lesmodels}}

This chapter lays the foundation for modeling the evolution of soot in large eddy simulation (LES) of turbulent nonpremixed combustion. In LES of turbulent flows, the large scale, geometry-dependent motions are resolved while small-scale, universal turbulent structures are modeled. The predictive capabilities of LES are dependent on the quality of the models used for the small-scale, subfilter phenomena. Accurately capturing the evolution of soot requires the integration of models for soot, turbulent combustion, and polycyclic aromatic hydrocarbons (PAH) into the LES framework.

The distribution of soot particle sizes, known as the Number Density Function (NDF), has been found to be bimodal due to the presence of small, nascent particles and larger, more mature aggregates. Modeling the NDF is achieved by first parameterizing it with the volume and surface area of particles. However, since the NDF is high-dimensional, the exact form cannot be solved for directly due to the associated computational costs. Thus, the Method of Moments is used to obtain the average quantities, known as moments, of the NDF. In this chapter, the transport equation that governs the evolution of these moments is introduced. This equation has a source term that requires modeling, so discussion of closure through the Hybrid Method of Moments is presented as well.

Turbulent combustion is modeled through the flamelet approach~\cite{peters1984}, where a three-dimensional turbulent flame is represented as an ensemble of locally one-dimensional laminar flamelets embedded in a turbulent flow field. This method relies on the separation of scales between the turbulence and the flame's reaction zone, allowing the thermochemical properties of the flame to be decoupled from the flow field. As a result, the flame structure can be evaluated \textit{a priori}, parameterized with a reduced set of variables, and stored in a database that is accessed during LES. In this chapter, modifications to the classical flamelet approach accounting for the formation of soot and radiative thermal losses are explained. Closure of subfilter quantities with the presumed PDF approach is also outlined.

Lastly, special consideration is given to gas-phase PAH. Since their chemistry is slow relative to the exothermic combustion chemistry, a transport equation is required to model the resulting unsteady effects. Discussion on modeling the source term for a lumped PAH species is included.

%% Using LES is advantageous because it captures unsteady flow features like swirl and recirculation that are neglected by other methods such as Reynolds-Averaged Navier-Stokes (RANS), without being as computationally intensive as full-fidelity Direct Numerical Simulation (DNS). Its predictive capabilities can lead to highly optimized designs through rapid iteration. The potential of LES to innovate the next generation of environmentally-friendly automobile or aircraft engines is the main reason it has been selected as the underlying modeling architecture.

\input{ch-lesmodels/soot}
\input{ch-lesmodels/combustion}
\input{ch-lesmodels/pdf}
\input{ch-lesmodels/pah}
