\section{Polycyclic Aromatic Hydrocarbons}
\label{sec:lesmodels:pah}

Understanding the formation and evolution of soot is central to this work, so accurately modeling the generation and consumption of gas-phase PAH that participate in nucleation and condensation is vital. A simple LES model for tracking the evolution of PAH would involve obtaining the PAH mass fraction from the RFPV combustion model. However, Attili \etal~\cite{attili2014} and Bisetti \etal~\cite{bisetti2012} noted in DNS studies of a nitrogen-diluted, \textit{n}-heptane/air turbulent nonpremixed flame that the PAH mass fraction deviates significantly from the values calculated by the steady flamelet approximation. The chemistry of PAH such as naphthalene is much slower than the chemistry of species that participate in the significantly exothermic reactions. As a result, PAH are unable to immediately adjust to the rapidly changing scalar dissipation rate field. They require a longer time to respond to the fluctuating field, so the unsteady term in the flamelet equations cannot be neglected.

To account for these unsteady phenomena, Mueller and Pitsch~\cite{mueller2012} developed a transport equation model for PAH that was inspired by a similar approach for slowly evolving $\ce{NO}$~\cite{ihme2008}. The spatially filtered transport equation is given by
\begin{equation}\label{eq:lesmodels:pah:pahtrans}
  \ftrder{\tf{Y}_{\text{PAH}}} = -\dsff[\tf{Y}_{\text{PAH}}]{\tf{u_j Y}_{\hspace{-3pt}\text{PAH}}} + \fdt[\text{PAH}]{\tf{Y}_{\text{PAH}}} + \fst[m]{\text{PAH}},
\end{equation}
where $\tf{Y}_{\text{PAH}}$ is the filtered mass fraction of a lumped PAH species and $\fst[m]{\text{PAH}}$ is the sum of the chemical source terms of all the PAH species. Under the assumption of unity Lewis number, all PAH species will have a diffusivity governed by $D_{\text{PAH}}$ and \cref{eq:lesmodels:pah:pahtrans} is the exact transport equation for a lumped PAH. However, if all PAH do not have the same molecular diffusivity, then the diffusivity of naphthalene is preferred for $D_{\text{PAH}}$ as it is generally the PAH species with the largest concentration. In this circumstance, \cref{eq:lesmodels:pah:pahtrans} is an approximate model for the evolution of the combined PAH mass fraction.

The unclosed terms in \cref{eq:lesmodels:pah:pahtrans} include the first and third terms on the right-hand side. The former is simply the subfilter flux and can be modeled using a conventional dynamic procedure as mentioned previously. Following Mueller and Pitsch~\cite{mueller2012}, the filtered chemical source term can be decomposed into a production term $\dot{m}_{+}$, a consumption term $\dot{m}_{-}$, and an additional consumption term due to the extraction of PAH from the gas-phase to form soot (dimerization) $\dot{m}_{D}$. For each individual PAH species $k$, the chemical production term is independent of $Y_k$, the chemical consumption term depends linearly on $Y_k$, and the dimerization term depends quadratically on $Y_k$. The source term of the lumped species will be decomposed similarly:
\begin{equation}\label{eq:lesmodels:pah:pahsrc}
  \fst[m]{\text{PAH}} = \mean{(\dot{m}_{+})} + \mean{\left( \frac{\dot{m}_{-}}{Y_{\text{PAH}}} \right) Y_{\text{PAH}}} + \mean{\left( \frac{\dot{m}_{D}}{{Y_{\text{PAH}}}^2} \right) {Y_{\text{PAH}}}^2},
\end{equation}
where the terms in parentheses are obtained directly from the RFPV combustion model because they are nearly independent of the PAH mass fraction. The last two terms on the right-hand side of \cref{eq:lesmodels:pah:pahsrc} are unclosed filtered correlations. Mueller and Pitsch~\cite{mueller2012} have proposed closure through a scale-similarity assumption between the transported PAH mass fraction and the PAH mass fraction computed from the steady flamelet equations. This model is given by
\begin{equation}\label{eq:lesmodels:pah:similarity}
  \frac{\tf{Y}_{\text{PAH}}}{\tf{Y}_{\text{PAH}}^{\text{RFPV}}} = \frac{Y_{\text{PAH}}^{''}}{Y_{\text{PAH}}^{\text{RFPV}''}},
\end{equation}
where the subscript $^{\text{RFPV}}$ designates quantities that are from the RFPV combustion model and the subscript $^{''}$ indicates the ``fluctuating'' LES quantities given by
\begin{equation}\label{eq:lesmodels:pah:residual}
  Y_{\text{PAH}}^{''} = Y_{\text{PAH}} - \tf{Y}_{\text{PAH}}.
\end{equation}
With this closure model, the filtered source term of the lumped PAH species is
\begin{equation}\label{eq:lesmodels:pah:pahsrcmodeled}
  \fst[m]{\text{PAH}} = \mean{\dot{m}}_{+}^{\text{RFPV}} + \mean{\dot{m}}_{-}^{\text{RFPV}}\left( \frac{\tf{Y}_{\text{PAH}}}{\tf{Y}_{\text{PAH}}^{\text{RFPV}}} \right) + \mean{\dot{m}}_{D}^{\text{RFPV}}\left( \frac{\tf{Y}_{\text{PAH}}}{\tf{Y}_{\text{PAH}}^{\text{RFPV}}} \right)^2.
\end{equation}
